The imprint of global magma oceans on exoplanet demographics

### 2022-12-15 src/tex/ms.tex
introduce hypothesis tests, form of hypothesis

\subsubsection{Definition of the magma ocean hypothesis and null hypothesis}
\todo{Define the null and alternative hypotheses}
\begin{note}
    As a null hypothesis, we consider the case where the planetary radius distribution is independent of the instellation,
    \begin{equation}
        h_{\mathrm{null}}(\theta, S) = \theta,
    \end{equation}
    where $\theta$ is the set of parameters defining the radius distribution.
    We further define an alternative hypothesis that describes radius changes due to steam atmospheres and water in the melt.
    As motivated above, this hypothesis takes the form of a step function in instellation $S$, where the step occurs at the outer edge of the runaway greenhouse region.
    Our main observable shall be the average planet radius in the planet population inside and outside this threshold.
    The magma ocean hypothesis is then defined as
\begin{equation}
    h_{\mathrm{MO}}(\theta, S) =
%        \begin{cases}
%        1 & \text{if } x \in \mathbb{Q}\\
%        0 & \text{if } x \in \mathbb{R}\setminus\mathbb{Q}
%    \end{cases}
        \begin{cases}
        h_{\mathrm{null}} & \text{for } S \leq S_\mathrm{thresh}\\
        <R_\mathrm{P}> (f_\mathrm{steam}, f_\mathrm{water}) & \text{for } S > S_\mathrm{thresh}.
    \end{cases}
\end{equation}
    Here, $f_\mathrm{steam}$ and $f_\mathrm{water}$ are predictions from the steam atmosphere and water incorporation models, respectively.
    They are assumed to act additively on the planet radii and thus on their average $<R_\mathrm{P}>$.

@@

For the likelihood function, we assumed that the planet radii $R_{\mathrm{P}, i}$ are measured with a normally distributed uncertainty $\sigma_{R_\mathrm{P}, i}$ and adopted a normal distribution
\begin{equation}
    \mathcal{L}(R_\mathrm{P} \mid \boldsymbol{\theta})=\prod_{i}^{N} \frac{1}{\sqrt{2 \pi \sigma_{R_\mathrm{P}, i}^{2}}} \exp \left(-\frac{\left(R_{\mathrm{P}, i}-h\left(\boldsymbol{\theta}, a_{\mathrm{eff}, i}\right)\right)^{2}}{2 \sigma_{R_\mathrm{P}, i}^{2}}\right).
\end{equation}
Here, $h\left(\boldsymbol{\theta}, a_{\mathrm{eff}, i}\right)$ corresponds to the functional form of the magma ocean hypothesis\todo{link to eqn defining the hypothesis} .
We tested this hypothesis against the null hypothesis $h_\mathrm{null} (\theta, R_\mathrm{P}) = \theta$, which states that there is no relationship between the measured planet radii and (scaled) semi-major axes.

@@

%\begin{acknowledgments}
%The authors thank Yann Fabel, Paul Molliere, and Terry-Ann Suer for insightful discussions.
%We are grateful to Martin Turbet for providing the mass-radius relationship data for planets harboring steam atmospheres.
%\end{acknowledgements}

>>>>

\subsubsection{Definition of the runaway greenhouse hypothesis}
We now turn to assessing the simulated surveys in terms of their ability to detect the runaway greenhouse transition and their constraining power on parameters of this trend.
To do this, we rely on a Bayesian hypothesis testing approach where we quantify the evidence of a hypothesis over another based on the (simulated) data.
For our specific problem, this implies comparing evidences for a demographic imprint of the runaway greenhouse effect to its absence.

\begin{note}
%    Bayesian model comparison requires that an alternative mode be specified against which the comparison is made.
\end{note}

\begin{note}
    As a null hypothesis, we consider the case where the planetary radius distribution is independent of the instellation,
    \begin{equation}
        h_{\mathrm{null}}(\theta, S) = \theta,
    \end{equation}
    where $\theta$ is the set of parameters defining the radius distribution.
    We further define an alternative hypothesis that describes radius changes due to steam atmospheres and water in the melt.
    As motivated above, this hypothesis takes the form of a step function in instellation $S$, where the step occurs at the outer edge of the runaway greenhouse region.
    Our main observable shall be the average planet radius in the planet population inside and outside this threshold.
    The magma ocean hypothesis is then defined as
\begin{equation}\label{eq:mo_hypo}
    h_{\mathrm{MO}}(\theta, S) =
%        \begin{cases}
%        1 & \text{if } x \in \mathbb{Q}\\
%        0 & \text{if } x \in \mathbb{R}\setminus\mathbb{Q}
%    \end{cases}
        \begin{cases}
        h_{\mathrm{null}} & \text{for } S \leq S_\mathrm{thresh}\\
        <R_\mathrm{P}> (f_\mathrm{steam}, f_\mathrm{water}) & \text{for } S > S_\mathrm{thresh}.
    \end{cases}
\end{equation}
    Here, $f_\mathrm{steam}$ and $f_\mathrm{water}$ are predictions from the steam atmosphere and water incorporation models, respectively.
    They are assumed to act additively on the planet radii and thus on their average $<R_\mathrm{P}>$.

@@

For the likelihood function, we assumed that the planet radii $R_{\mathrm{P}, i}$ are measured with a normally distributed uncertainty $\sigma_{R_\mathrm{P}, i}$ and adopted a normal distribution
\begin{equation}
    \mathcal{L}(R_\mathrm{P} \mid \boldsymbol{\theta})=\prod_{i}^{N} \frac{1}{\sqrt{2 \pi \sigma_{R_\mathrm{P}, i}^{2}}} \exp \left(-\frac{\left(R_{\mathrm{P}, i}-h\left(\boldsymbol{\theta}, S_i\right)\right)^{2}}{2 \sigma_{R_\mathrm{P}, i}^{2}}\right).
\end{equation}
Here, $h\left(\boldsymbol{\theta}, S_i\right)$ corresponds to the functional form of the magma ocean hypothesis~(Eqn.~\ref{eq:mo_hypo}).
We tested this hypothesis against the null hypothesis $h_\mathrm{null} (\theta, R_\mathrm{P}) = \theta$, which states that there is no relationship between the measured planet radii and received instellation.

@@

%\begin{acknowledgments}
%The authors thank Yann Fabel, Paul Molliere, Matthew Murphy, and Terry-Ann Suer for insightful discussions.
%We are grateful to Martin Turbet for providing the mass-radius relationship data for planets harboring steam atmospheres.
%\end{acknowledgements}
